Reassessing the role of theory and machine learning in forecasting civil conflict

### 2020-09-18 journal-survey/README.Rmd
finish data entry for B&S references survey

@@

# take out items that are not coded yet (missing binary_outcome column)
articles <- articles[!is.na(articles$binary_outcome), ]

# if one of the payload columns is NA, set it to 0
articles$uses_auc[is.na(articles$uses_auc)] <- 0L
articles$uses_roc[is.na(articles$uses_roc)] <- 0L
articles$uses_smoothed_auc[is.na(articles$uses_smoothed_auc)] <- 0L
articles$uses_smoothed_roc[is.na(articles$uses_smoothed_roc)] <- 0L

@@

smry <- binary %>%
  mutate(any_smoothed = uses_smoothed_auc | uses_smoothed_roc,
         any_empirical = uses_auc | uses_roc,
         uses_any = any_smoothed | any_empirical,
         only_smoothed = any_smoothed & !any_empirical,
         only_empirical = any_empirical & !any_smoothed,
         both = any_smoothed & any_empirical) %>%
  summarize(
    articles_coded = nrow(articles),
    articles_with_binary_outcome = n(),
    use_any_roc_measure = sum(uses_any),
    only_smoothed = sum(only_smoothed),
    only_empirical = sum(only_empirical),
    both = sum(both)
  ) %>%
  pivot_longer(everything())

@@

>>>>

The sample consists of:

1. All references in the Blair and Sambanis paper, excluding books and non-academic articles. 
2. Two issues of JCR; 2020 volumes 1 and 9.

@@

# take out items that are not coded yet (missing "journal" column)
articles <- articles[!is.na(articles$journal), ]

# if one of the payload columns is NA, set it to 0
articles$uses_auc[is.na(articles$uses_auc)] <- 0L
articles$uses_roc[is.na(articles$uses_roc)] <- 0L
articles$uses_smoothed_auc[is.na(articles$uses_smoothed_auc)] <- 0L
articles$uses_smoothed_roc[is.na(articles$uses_smoothed_roc)] <- 0L
articles$unclear_auc[is.na(articles$unclear_auc)] <- 0L
articles$unclear_roc[is.na(articles$unclear_roc)] <- 0L

@@

binary <- binary %>%
  mutate(any_smoothed = uses_smoothed_auc | uses_smoothed_roc,
         any_empirical = uses_auc | uses_roc,
         any_unclear = unclear_roc | unclear_auc,
         uses_any = any_smoothed | any_empirical | any_unclear,
         only_smoothed = any_smoothed & !any_empirical & !any_unclear,
         only_empirical = any_empirical & !any_smoothed & !any_unclear,
         only_unclear = any_unclear & !any_empirical & !any_smoothed)

smry <- binary %>%
  summarize(
    items_coded = nrow(articles),
    articles_with_binary_outcome = n(),
    use_any_roc_measure = sum(uses_any),
    only_smoothed = sum(only_smoothed),
    only_empirical = sum(only_empirical),
    only_unclear = sum(only_unclear)
  ) %>%
  pivot_longer(everything())

@@

What items had unclear or ambiguous codings?

```{r}
binary %>% 
  filter(any_unclear) %>%
  select(title, year, journal, notes) %>%
  knitr::kable()
```

Note that both articles include Blair as co-author.

## Data

- "uses_roc", "uses_smoothed_roc", "unclear_roc": Does the paper include figures showing (empirical / smoothed / unclear) ROC curves?
- "uses_auc", "uses_smoothed_auc", "unclear_auc": Does the paper include tables showing (empirical / smoothed / unclear) AUC-ROC values?

If the paper includes visually obvious empirical ROC curves, we assumed that any corresponding AUC-ROC values were also generated from empirical ROC curves.

Some papers had AUC values but no ROC curves. In these cases we consulted replication code. These are listed below.

```{r}
articles %>% 
  filter(!uses_roc & uses_auc) %>%
  select(title, year, journal, notes) %>%
  knitr::kable()
```

```{r}
write_csv(articles, "coded-items.csv")
```

### 2020-09-18 paper.Rmd
finish data entry for B&S references survey

The most consequential issue that we found is that all AUC-ROC values reported in B&S Tables 1 and 2 are calculated using smoothed ROC curves, not the original, empirical ROC curves. The data in rare events problems like this one restrict the number of true positive rate values and lead to non-continuous ROC curves; B&S refer to smoothing in the context of their Figure 1 with ROC curves justified as easing interpretation. Smoothing is used to produce all AUC-ROC values they report. There is no difficulty to be overcome in using non-smoothed plots. Indeed, such plots are standard practice in conflict research and forecasting, both for visualization and when calculating AUCs. We examined N articles in K journals over the period from 20xx-2020 looking for articles which have a binary dependent variable and use the ROC techonology in the analysis section. We found that XX percent of these articles use the empirical ROC calculations and only 100-XX use a smoothed curve.  This is not surprising since the original ideal as well as contemporaneous documentation suggest this is the gold standard.  What is surprising is that anyone would choose the smoothed ROC implementation.
INSERT SMALL STUDY HERE WITH DETAILS.

>>>>

The most consequential issue that we found is that all AUC-ROC values reported in B&S Tables 1 and 2 are calculated using smoothed ROC curves, not the original, empirical ROC curves. The data in rare events problems like this one restrict the number of true positive rate values and lead to non-continuous ROC curves; B&S refer to smoothing in the context of their Figure 1 with ROC curves justified as easing interpretation. Smoothing is used to produce all AUC-ROC values they report. There is no difficulty to be overcome in using non-smoothed plots. Indeed, such plots are standard practice in conflict research and forecasting, both for visualization and when calculating AUCs. We examined all 63 references in B&S and also 15 articles in two randomly picked issues of JCR for 2020 (numbers 1 and 9), for a total of 78 items, in order to survey the use of ROC curves. Of the 37 articles that used a binary outcome, 19 included either ROC curve figures or a table with AUC-ROC values. We found that 90%--17 of 19--clearly used empirical ROC curves, and 2 (10%) where we could not establish whether empirical or smoothed curves were used.\footnote{One paper reports AUC-ROC values but no figure with ROC curves, and we could not find publicly posted replication code. The other is an ambiguous case where ROC curves are manually constructed in a way that might be considered "smoothing"--and one of the two figures certainly looks smoothed--but it's not a parametric smoothing method like used in B&S. The online replication archive has more details at https://github.com/andybega/Blair-Sambanis-replication/tree/master/journal-survey.} (Both of the ambiguous cases include Blair as coauthor.) The survey results are not surprising since the original ideal as well as contemporaneous documentation suggest this is the gold standard.  What is surprising is that anyone would choose the smoothed ROC implementation.

INSERT SMALL STUDY HERE WITH DETAILS.
